Causal Compatibility Inequalities Admitting of Quantum Violations in the Triangle Scenario

### 2017-01-04 documents/437Slides/slides.tex
Yevick suggestions

\begin{frame}
    \frametitle{This Talk}
    \begin{enumerate}
        \item Motive why quantum-classical difference is worth studying
        \item Explain why the triangle scenario is interesting
        \item Briefly discuss how we obtained new measures of non-classicality
        \item Discuss search for new non-classical quantum correlations
    \end{enumerate}
\end{frame}

@@

\begin{frame}
    \frametitle{Notation \& Jargon}
    \begin{center}
        Which observations are \textit{compatible} with a given classical model?
    \end{center}
    \begin{columns}
        \column{0.45\textwidth}
        \begin{center}
            \textbf{Causal Structure $\graph$}\\
            {\footnotesize (classical model)} \\
            \[ \graph = \br{\nodes, \edges} \]
            \vspace{-0.3in}
            \begin{align*}
                \edges &: \text{\term{causal influence}} \\
                \nodes &: \text{\term{random variables}} \\
                \nodes_{O} &: \text{\term{observable variables}} \\
                \nodes_{L} &: \text{\term{latent variables}}
            \end{align*}
            \[ \nodes = \nodes_{O} \cup \nodes_{L} \]
        \end{center}
        \column{0.45\textwidth}
        \begin{center}
            \textbf{Marginal Model $\prob^{\mscenario}$} \\
            {\footnotesize (observations)}
            \[ \mscenario : \text{\term{marginal scenario}} \]
            \[ \mscenario = \bc{V \mid V \subseteq \nodes_{O}} \]
            \[ \prob^{\mscenario} = \bc{\prob[V] \mid V \in \mscenario} \]
        \end{center}
    \end{columns}
\end{frame}

\begin{frame}
    \frametitle{The Triangle Scenario}
    \begin{center}
        \scalebox{1.5}{\input{../../figures/causal_structures/triangle_scenario}}
    \end{center}
\end{frame}

\begin{frame}
    \frametitle{Causal Compatibility}
    \textbf{Question:} Is $\prob^{\mscenario}$ \term{compatible} with $\graph$?\\
    \textbf{Answer:} If $\prob^{\mscenario}$ is to be compatible with $\graph$ then:
    \begin{enumerate}
        \item $\prob^{\mscenario}$ must admit a \term{joint distribution} $\prob[\nodes]$:
        \[ \prob[V] = \sum_{\nodes \setminus V} \prob[\nodes] \]
        \item $\prob[\nodes]$ must obey \term{Markov conditions}:
        \[ \prob[\nodes] = \prod_{n\in\nodes}\prob[n \mid \Pa[\graph]{n}] \]
    \end{enumerate}
    \textbf{Inequality:} A \term{casual compatibility inequality} $I$ is an inequality over $\prob^{\mscenario}$ that is satisfied by all compatible $\prob^{\mscenario}$
\end{frame}

@@

\begin{frame}
    \frametitle{Why Triangle Scenario?}
    \begin{itemize}
        \item Triangle scenario (TS) has be studied extensively before
        \item In~\cite{Branciard_2012} it was noted that characterizing quantum non-classicality in TS remained an \textbf{open problem} and that identifying compatibility inequalities \textbf{seemed challenging}
        \item In~\cite{Fritz_2012}, Fritz demonstrated that TS is the \textbf{smallest} correlation scenario in which quantum non-classicality is manifest (proof without inequalities)
        \item In~\cite{Henson_2014}, TS was classified as an \textbf{interesting} causal structure: conditional independence relations are not a sufficient characterization of compatibility (there are none)
        \item Several other authors (see~\cite{Steudel_2010},~\cite{Chaves_2014},~\cite{Inflation}, $\ldots$) have been unable to find inequalities with quantum violations
    \end{itemize}
\end{frame}

>>>>

% \begin{frame}
%     \frametitle{This Talk}
%     \begin{enumerate}
%         \item Motive why quantum-classical difference is worth studying
%         \item Explain why the triangle scenario is interesting
%         \item Briefly discuss how we obtained new measures of non-classicality
%         \item Discuss search for new non-classical quantum correlations
%     \end{enumerate}
% \end{frame}

@@

\begin{frame}
    \frametitle{Notation \& Jargon}
    \begin{center}
        Given a set of observations, are they \textit{compatible} with a given classical model?
    \end{center}
    \begin{columns}
        \column{0.45\textwidth}
        \begin{center}
            \textbf{Causal Structure $\graph$}\\
            {\footnotesize (classical model)} \\
            \[ \graph = \br{\nodes, \edges} \]
            \vspace{-0.3in}
            \begin{align*}
                \edges &: \text{\term{causal influence}} \\
                \nodes &: \text{\term{random variables}} \\
                \nodes_{O} &: \text{\term{observable variables}} \\
                \nodes_{L} &: \text{\term{latent variables}}
            \end{align*}
            \[ \nodes = \nodes_{O} \cup \nodes_{L} \]
        \end{center}
        \column{0.45\textwidth}
        \begin{center}
            \textbf{Marginal Model $\prob^{\mscenario}$} \\
            {\footnotesize (set of observations)}
            \[ \mscenario : \text{\term{marginal scenario}} \]
            \[ \mscenario = \bc{V \mid V \subseteq \nodes_{O}} \]
            \[ \prob^{\mscenario} = \bc{\prob[V] \mid V \in \mscenario} \]
        \end{center}
    \end{columns}
\end{frame}

\begin{frame}
    \frametitle{Causal Compatibility}
    \textbf{Question:} Can your model $\graph$ explain the observation $\prob^{\mscenario}$?\\
    \textbf{Question:} Is $\prob^{\mscenario}$ \term{compatible} with $\graph$?\\
    \vspace{0.3in}
    \textbf{Answer:} If $\prob^{\mscenario}$ is to be compatible with $\graph$ then:
    \begin{enumerate}
        \item $\prob^{\mscenario}$ must admit a \term{joint distribution} $\prob[\nodes]$:
        \[ \prob[V] = \sum_{\nodes \setminus V} \prob[\nodes] \]
        \item $\prob[\nodes]$ must obey \term{Markov conditions} (independence given parents):
        \[ \prob[\nodes] = \prod_{n\in\nodes}\prob[n \mid \Pa[\graph]{n}] \]
    \end{enumerate}
\end{frame}

\begin{frame}
    \frametitle{Inequalities}
    \textbf{Task:} Determine if $\prob^{\mscenario}$ is \term{compatible} or \term{incompatible} with $\graph$\\
    \vspace{0.3in}
    \textbf{Tool:} A \term{casual compatibility inequality} $I$ is an inequality satisfied by all compatible $\prob^{\mscenario}$
    \vspace{0.3in}
    \begin{itemize}
        \item Inequalities $I$ are \textbf{extremely useful}:
        \item If $\prob^{\mscenario}$ satisfies \textit{all} such constraints $I$, then it is \textit{compatible} with $\graph$
        \item If $\prob^{\mscenario}$ violates \textit{any} constraint $I$, then it is \textit{incompatible} with $\graph$
        \item Inequalities act as \textit{measures} of non-classicality
        \item $I$ identifies \textit{fundamentally} new resources or quantum advantages!
    \end{itemize}
\end{frame}

\begin{frame}
    \frametitle{Case Study: The Triangle Scenario}
    \begin{center}
        \scalebox{1.5}{\input{../../figures/causal_structures/triangle_scenario}}
    \end{center}
\end{frame}

@@

\begin{frame}
    \frametitle{Why Triangle Scenario?}
    \begin{itemize}
        \item The TS could be home to \textbf{brand new} entanglement resources which could lead to new technologies and fundamental physics, albeit very difficult to find
        \item In~\cite{Branciard_2012} it was noted that characterizing quantum non-classicality in TS remained an \textbf{open problem} and that identifying compatibility inequalities \textbf{seemed challenging}
        \item In~\cite{Fritz_2012}, Fritz demonstrated that TS is the \textbf{smallest} correlation scenario in which quantum non-classicality is manifest (proof without inequalities)
        \item In~\cite{Henson_2014}, TS was classified as an \textbf{interesting} causal structure: conditional independence relations are not a sufficient characterization of compatibility (there are none)
        \item Several other authors (see~\cite{Steudel_2010},~\cite{Chaves_2014},~\cite{Inflation}, $\ldots$) have been unable to find inequalities with quantum violations
    \end{itemize}
\end{frame}
